Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections

### 2025-03-19 docs/index.qmd
clean repo and push to publish

---
title: title
subtitle: subtitle
shorttitle: A Short title # aog-article-pdata
shortauthors: Isaiah      # aog-article-pdata
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
date: "2024-10-16"
date-modified: today
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
editor_options: 
  chunk_output_type: console
---

@@

# load packages
library(tidyverse)
library(easystats)
library(odds.n.ends) # model sig from glm
library(patchwork)
library(texreg) # Conversion of R regression output into latex or HTML tables

@@

# load data dictionary/codebook
codebook <- read.csv("data/survey_codebook.csv")
data_dict <- read.csv("data/data_dictionary.csv")
```

```{r}
#| label: setup-theme

# set ggplot theme that will be used for every plot
# theme_get()
old <- ggplot2::theme_set(ggplot2::theme_bw())  # capture current theme

# set/update new theme
ggplot2::theme_set(
  new = ggplot2::theme_bw(
    base_family = 'TeX Gyre Pagella')+ # set font of ggplots
    theme(legend.position = 'bottom')  # set legend position for all plots
)

@@

In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned[^3]. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate.

@@

Intuitively, enhancing public trust in elections would best be accomplished by enhancing the *trustworthiness* of the institution, i.e., consistently doing the things that election officials already regularly do come election time. However, trust and confidence in elections has become ever more precarious over the last few election cycles. Especially considering public polling data since 2000 shows that confidence that votes were, or would be, counted as intended was in a consistent decline despite efforts towards bolstering election integrity and trustworthiness [@sances2015]. This is even more pronounced considering the role that partisanship has had on such confidence over accuracy of vote count [@sances2015; @stewart2022].

@@

As such, it is questionable whether we can characterize public trust in the pre-election period as the same trust after the election results have come out. The former is *anticipatory*---i.e., the kind that is more or less anxious given the uncertainties surrounding the election. The latter is *empirical*---a judgement discerned in hindsight after the experience of the election event has occurred. As discussed by @stewart2022, not only was confidence in vote count accuracy influenced by the voter experience, but so too were evaluations of election officials [-@stewart2022, 242-243]. Not to mention the influence that the election results would also have on such evaluations. In other words, trust evaluations post-election are influenced by one's interpretation of their experience and election results. This study, in contrast, focuses primarily on that *anticipatory* kind of confidence, which speaks more to those insecurities based on perceptions of the institution's trustworthiness than upon the particular voter experience.

@@

Election officials are likely to be agnostic as to who dedicates their time to civil service such as election work. Staffing issues have been an issue since at least the 1990s [@maidenberg1996; @ferrer2024]. In 2020, such efforts were made far more difficult by the COVID-19 pandemic [@mena2020; @abbate2020a]. There's no special reason to target veterans for recruitment above other groups. Indeed, there's no reason to discriminate recruitment efforts at all if the point is purely to fill staffing vacancies.

@@

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve confidence in elections and ease insecurity, a recent experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment ($n =$ 650) control ($n =$ 637) conditions[^5].

Survey participants read either a treatment or control vignette, which resembled a news article about efforts in Maricopa County, AZ to recruit election staff and volunteers for the 2024 general election[^6]. The treatment vignette referred to a program designed to recruit veterans and their family members and describes an interviewee "Jordan Braxton" as an Army veteran. The control vignette simply omitted any mention of veterans and their family members, and didn't describe "Jordan Braxton" as an Army veteran. Beyond those small differences and the headlines, the article vignettes are identical. Therefore, effects can be attributed to the information about veterans in the treatment vignette[^7].

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignette due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. Because of this, treatment effects are potentially limited or constrained to attitudes concerning the location specified in the vignette. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. Later, I compare and discuss results between the items pertaining to different settings.

@@

The items in this study all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, the survey of this study focused on attitudes regarding election workers such as election officials, staff, and volunteers specific to the location mentioned in the vignette---Maricopa County, AZ---as well as specific to one's local area[^8].

The survey included items that assessed both trust in elections administration and expectations of electoral fraud, referred to simply as *trust* and *distrust*, respectively. Responses from these two sets of items were utilized to construct the primary dependent variable, *confidence in elections*.

Prior to the reading the treatment or control vignette, participants responded to distinct survey items assessing one's favorability to local election officials ("In general, how favorable or unfavorable is your impression of local election officials?") and legitimacy of the 2020 election results ("Regardless of whom you supported in the 2020 election, do you think Joe Biden's election as president was legitimate, or was he not legitimately elected?"). The former presented five response options from "Strongly favorable" to "Strongly unfavorable"; the latter item presented a binary of either "Legitimate" or "Not Legitimate".

@@

Finally, a pair of survey items assessed concerns for potential violence and confidence in voter safety. The pair was asked once in regard to Maricopa County, AZ and again with respect to one's local area[^9].

@@

kableExtra::kbl(conf_items_text, booktabs = T,
                col.names = c("Trust", "Distrust"),
                linesep = "\\addlinespace") |> 
  kableExtra::kable_styling(full_width = T,
                            font_size = 8,
                            latex_options = c("HOLD_position")) |> 
  kableExtra::column_spec(column = 1:2, width = "8cm")

@@

```{r}
#| label: tbl-ttest
#| tbl-cap: "Mean Difference of Treatment Compared to Control Condition"
#| tbl-pos: H

# conduct two t-tests using `rstatix::t_test`, bind results by row, and pipe
# into table
dplyr::bind_rows(
  data |> mutate(group = forcats::fct_relevel(group, "Treatment", after = 0L)) |>
  rstatix::t_test(formula = conf.az.rescaled ~ group,var.equal = F,detailed = T),
  data |> mutate(group = forcats::fct_relevel(group, "Treatment", after = 0L)) |>
  rstatix::t_test(formula = conf.lcl.rescaled ~ group,var.equal = F, detailed = T)
  ) |> 
  # specify columns and column names
  select(
    Diff = estimate,
    treat_mean = estimate1,
    control_mean = estimate2,
    treat_n = n1,
    control_n = n2,
    t = statistic,
    p, df, conf.low, conf.high) |> 
  mutate(
    Place = dplyr::case_when(
      p <= 0.05 ~ "Maricopa County, AZ", 
      p > 0.05 ~ "Local Area"), 
    .before = Diff) |> 
  insight::format_table(stars = T) |> 
  # insight::export_table(format = "md") # alternate
  # output table as `kable` for LaTeX pdata
  kableExtra::kbl( 
    booktabs = T,
    col.names = c("Place", "$\\bar{x}_{diff}$", "$\\bar{x}_{treat}$", "$\\bar{x}_{control}$",
                  "$treat_{n}$", "$control_{n}$", "t", "p", "data", "CI"), 
    escape = F) |> 
  kableExtra::footnote(general = "Welch Two Sample t-tests of difference between Confidence in Elections by Treatment Condition.") |> 
  kableExtra::kable_styling(
    font_size = 9, 
    protect_latex = T)

@@

::: {.content-visible when-format="bookup-html" unless-format="submittable-pdf"}

@@

::: {.content-visible when-format="submittable-pdf" unless-format="bookup-html"}

@@

@@

Publicized announcements that election officials are actively recruiting former military service members to work as election staff and volunteers is associated with slightly greater confidence in elections compared to election worker recruitment announcements that do not explicitly mention targeted recruitment of former military service members. Moreover, increased levels of confidence is limited to elections beyond one's local area and, of course, in anticipation of election night. Whether the increase in confidence is sustained post-election night cannot be inferred from these results.

Bear in mind that confidence in elections for one's local area is already higher than for those beyond. Thus, there's already more insecurity concerning elections elsewhere for whatever reason. In the anticipatory period prior to national elections, publicized efforts to recruit veterans to work as election staff and volunteers may be a small, but positive, step towards reducing insecurity in elections that occur elsewhere. Especially among those who maintained a steadataast degree of distrust in elections. Despite a lack of evidence that systematic electoral fraud had occurred in the 2020 election, a substantial proportion of the population were unwilling to affirm the legitimacy of the results. However, despite such legitimacy beliefs, most were not consistent as they expressed varying levels of trust in different aspects of elections administration.

These results speaks more to the influence that military veterans have among those who distrust the electoral process and refute the legitimacy of prior outcomes than it does to the public more generally. In general, public confidence appears strongest when the elections under consideration regard those within one's local community. Yet, such strong confidence is not generalized outward to the institution of elections administration as a whole. Insecurity in the integrity of the electoral process and its administration may be naturally higher for elections that occur further from home. Whether that is due to some local favorability bias or some outward negativity bias is hard to say. And seeing as how insecurity presents vulnerability, it is perhaps far easier to inspire a general sense of distrust in the process and its administration in places not near and around one's local area.

@@

## Survey Items

```{r}
#| label: tbl-4
#| tbl-cap: "Survey Items and Response Options"

data |> 
  select(q6, q7, q25, q26) |> 
  labelled::generate_dictionary() |> 
  dplyr::as_tibble() |> 
  mutate(value_labels = dplyr::case_when(
    variable == 'q6' ~ "[Strongly unfavorable, Somewhat unfavorable, Neither, Somewhat favorable, Strongly favorable]",
    variable == 'q7' ~ "Legitimate, Not legitimate",
    variable == 'q25' ~ "Very concerned, Somewhat concerned, Not too concerned, Not at all concerned",
    variable == 'q26' ~ "Not at all confident, Not too confident, Somewhat confident, Very confident"
  )) |> 
  select(Question = label, response = value_labels) |> 
  kableExtra::kbl(booktabs = T, linesep = "\\addlinespace") |> 
  kableExtra::kable_styling(full_width = T,
                            font_size = 8,
                            latex_options = c("HOLD_position")) |> 
  kableExtra::column_spec(column = 1, width = "7cm") |> 
  kableExtra::column_spec(column = 2, width = "4cm")

@@

@@

```{r}
#| label: tbl-demog
#| tbl-cap: "Sample Demographics"
#|

@@

@tbl-logit shows results of a logistic regression test of random assignment to the treatment group. Demographics such as age, gender, race, educational attainment, and party ID are included as predictor variables. Note that 19 missing observations were deleted. A $\chi^2(15) = 11.010$, with 15 degrees of freedom and associated p-value > 0.05 ($p = 0.75$) confirms that none of the demographic predictor variables significantly increased the log-odds---in turn, the probability---of being assigned to the treatment group.

@@

# simple logit of test of random assignment to treatment
# treatment is DV (treatment = 1, control = 0)
# demographic variables are IVs

data <- data |> 
  mutate(group.dum = dplyr::case_when(
    group == "Treatment" ~ 1,
    group == "Control" ~ 0,
    TRUE ~ NA), .after = group)

m1 <- glm(group.dum ~ age_4cat + gender_3cat + race2 + educ_4cat + partyid_3cat, 
          family = binomial(link = "logit"), data = data)

# A Chi-square of 11.01 with 15 degrees of freedom and associated p-value > 0.05 (p-value = 0.752) confirms that none of the demographic predictor variables significantly increased the probability of being assigned to the treatment group.

logit_tab <- gtsummary::tbl_regression(
    m1, 
    intercept = T,
    label = list(
      age_4cat ~"Age", 
      gender_3cat ~ "Gender", 
      race2 ~ "Race (includes Hispanic as race)",
      educ_4cat ~ "Educational Attainment", 
      partyid_3cat ~ "Party ID")) |> 
  gtsummary::add_global_p(keep = T) |> 
  gtsummary::add_nevent(location = c("label", "level")) |>
  gtsummary::bold_p() |> 
  gtsummary::bold_labels() |>
  # gtsummary::show_header_names()
  gtsummary::modify_header(
    label = " ",
    stat_nevent = "**Treatment N**") 
# 
#   gtsummary::modify_footnote_header(
#     footnote = "Chi-Squared ($\\chi^2$) = 11.01, d.f. = 15, p = 0.75", columns = "p.value")

gtsummary::as_kable_extra(logit_tab, 
                          booktabs = T, 
                          escape = F, 
                          addtl_fmt = T) |> 
  kableExtra::footnote("Chi-Squared = 11.01, d.f. = 15, p = 0.75", escape = F) |> 
  kableExtra::kable_styling(font_size = 9)

@@

@@

>>>>

---
title: 'Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
# subtitle: 'The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections'
shorttitle: 'Effect of Recruiting Military Veterans on Confidence in 2024 US Elections' # aog-article-pdf
shortauthors: Isaiah      # aog-article-pdf
abstract: |
  In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate. In addition, concerns over voter safety and potential for violence were lower among those who received the treatment.
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
editor_options: 
  chunk_output_type: console
---

@@

# load packages
library(tidyverse)
library(modelsummary)
library(tinytable) # simple tables in html, LaTeX, markdown, word, pdf, typst
library(litedown)
library(ordinal)
library(marginaleffects)
library(easystats)
library(patchwork)

# set some session options for `tinytable`
options(tinytable_tt_digits = 3)
# options(tinytable_theme_placement_latex_float = "H")
options(tinytable_html_mathjax = TRUE) # set option to render math in html

@@

@@

In this paper, I report results from a recent survey experiment administered to test whether publicized efforts to recruit veterans to work as election staff and volunteers would improve public trust in elections and ease election insecurity. Results of the survey experiment support the notion that emphasizing veterans as the target of election worker recruitment efforts eases pre-election insecurity. Expectations of electoral fraud and concerns for voter safety were lower among those who read an announcement that veterans are being recruited to work as election staff and volunteers compared to those who read a control vignette where veterans were not mentioned[^2]. Notable is that there was a significant difference in confidence among those in the treatment condition who believe that results of the 2020 election were illegitimate. In addition, concerns over voter safety and potential for violence were lower among those who received the treatment.

@@

Intuitively, enhancing public trust in elections would best be accomplished by enhancing the *trustworthiness* of the institution, e.g., enhancing vote tabulation equipment and technology, conducting audits. However, trust and confidence in elections has become ever more precarious over the last few election cycles. Especially considering public polling data since 2000 shows that confidence that votes were, or would be, counted as intended was in a consistent decline despite efforts towards bolstering election integrity and trustworthiness [@sances2015]. This is even more pronounced considering the role that partisanship has had on such confidence over accuracy of vote counts [@sances2015; @stewart2022].

@@

As such, it is questionable whether we can characterize public trust in the pre-election period as the same trust after the election results have come out. The former is *anticipatory*---i.e., the kind that is more or less anxious given the uncertainties surrounding the election. The latter is *empirical*---a judgement discerned in hindsight after the experience of the election event has occurred. As discussed by @stewart2022, not only was confidence in vote count accuracy influenced by the voter experience, but so too were evaluations of election officials [-@stewart2022, 242-243]. Not to mention the influence that the election results would also have on such evaluations. In other words, evaluations of trust considered in hindsight are influenced by the voter experience and the outcome of the election results. This study, in contrast, focuses primarily on that *anticipatory* kind of confidence, which speaks more to those insecurities based on perceptions of the institution's trustworthiness than upon the particular voter experience.

@@

Election officials are likely to be agnostic as to who dedicates their time to civil service such as election work. Especially since staffing issues have been an issue since at least the 1990s [@maidenberg1996; @ferrer2024]. In 2020, such efforts were made far more difficult by the COVID-19 pandemic [@mena2020; @abbate2020a]. There's no special reason to target veterans for recruitment above other groups. Indeed, there's no reason to discriminate recruitment efforts at all if the point is purely to fill staffing vacancies.

@@

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve confidence in elections and ease insecurity, a recent experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment ($n =$ 650) control ($n =$ 637) conditions[^4].

Survey participants read either a treatment or control vignette, which resembled a news article about efforts in Maricopa County, AZ to recruit election staff and volunteers for the 2024 general election[^5]. The treatment vignette referred to a program designed to recruit veterans and their family members and describes an interviewee "Jordan Braxton" as an Army veteran. The control vignette simply omitted any mention of veterans and their family members, and didn't describe "Jordan Braxton" as an Army veteran. Beyond those small differences and the headlines, the article vignettes are identical. Therefore, effects can be attributed to the information about veterans in the treatment vignette[^6].

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignette due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. Because of this, treatment effects are potentially limited or constrained to attitudes concerning the location specified in the vignette. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. I compare and discuss results between the items pertaining to different settings.

@@

The items in this study all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, the survey of this study focused on attitudes regarding election workers such as election officials, staff, and volunteers specific to the location mentioned in the vignette---Maricopa County, AZ---as well as specific to one's local area[^7].

The survey included items that assessed both trust in elections administration and expectations of electoral fraud, referred to simply as *trust* and *distrust*, respectively. Responses from these two sets of items were utilized to construct the primary dependent variable, *confidence in elections*. Additionally, a pair of survey items assessed concerns for potential violence and confidence in voter safety. The pair was asked once in regard to Maricopa County, AZ and again with respect to one's local area[^8]. Although two survey items are insufficient to serve as a composite scale, safety concerns will be evaluated by examination of response to these two questions in particular.

Prior to the reading the treatment or control vignette, participants responded to distinct survey items assessing one's beliefs about the legitimacy of the 2020 election results ("Regardless of whom you supported in the 2020 election, do you think Joe Biden's election as president was legitimate, or was he not legitimately elected?"). Respondents selected either "Legitimate" or "Not Legitimate".

@@

@@

tinytable::tt(conf_items_text, width = c(.5, .5)) |> 
  tinytable::style_tt(fontsize = 0.80)

@@

```{r}
#| label: tbl-ttest
#| tbl-cap: "Mean Difference of Treatment Compared to Control Condition"
#| tbl-colwidths: 80
#| tbl-pos: H

# conduct two t-tests using `rstatix::t_test`, bind results by row, and pipe
# into table
tab_ttest <- dplyr::bind_rows(
  data |> mutate(group = forcats::fct_relevel(group, "Treatment", after = 0L)) |>
  rstatix::t_test(formula = conf.az.rescaled ~ group,var.equal = F,detailed = T),
  data |> mutate(group = forcats::fct_relevel(group, "Treatment", after = 0L)) |>
  rstatix::t_test(formula = conf.lcl.rescaled ~ group,var.equal = F, detailed = T)
  ) |> 
  # specify columns and column names
  select(
    Diff = estimate,
    treat_mean = estimate1,
    control_mean = estimate2,
    treat_n = n1,
    control_n = n2,
    t = statistic,
    p, df, conf.low, conf.high) |> 
  mutate(
    Place = dplyr::case_when(
      p <= 0.05 ~ "Maricopa County, AZ", 
      p > 0.05 ~ "Local Area"), 
    .before = Diff) |> 
  insight::format_table(stars = T)

# Output using `tinytable` R Package
colnames(tab_ttest) <- c(
  "Place",
  "$\\bar{x}_{diff}$",
  "$\\bar{x}_{treat}$",
  "$\\bar{x}_{control}$",
  "$n_{treat}$",
  "$n_{control}$",
  "t",
  "p",
  "df",
  "CI")

tinytable::tt(tab_ttest) |>
  tinytable::format_tt(
    j = 2:6, # specify columns  
    math = T # apply mathmode to specified columns
  ) |> 
  tinytable::style_tt(fontsize = 0.75) # fontsize in em units (0.75*default fontsize)

@@

::: {.content-visible when-format="bookup-html" unless-format="pdf"}

@@

::: {.content-visible when-format="pdf" unless-format="bookup-html"}

@@

### Concerns for Violence and Voter Safety

Thus far, publicized announcements that election officials will be targeting veterans in their recruitment efforts does quite a bit to improve confidence in elections outside of one's local area (or, at least for elections held in Maricopa County, AZ), especially among those who believed the 2020 election results were illegitimate. While distrust is one concern during the pre-election period, it is not the only concern that arises in anticipation of US elections; the public must feel confident that voters are safe to cast a ballot in person. The survey included questions^[See @tbl-4 in Appendix A] that allows me to assess whether those in the treatment condition expressed less concern for violence and more confidence in voter safety compared to those in the control condition. 
```{r}
#| label: fig-safety-bars
#| layout-ncol: 1
#| layout-nrow: 2
#| fig-cap: "Safety Concerns and Confidence by Treatment Condition"
#| fig-subcap: 
#|   - "Concern for Violence while voting in Maricopa County, AZ Elections"
#|   - "Confidence in Voter Saftey at Election Sites in Maricopa County, AZ"

# concern for violence by treatment
bar_q25 <- custom_barplot(
  data = data, 
  x = q25, 
  group = group,
  subtitle = str_wrap(attr(data$q25, "label"), width = 85)
  )+
  theme(text = element_text(family = 'TeX Gyre Pagella'),
        legend.position = "right")

# confidence in voter safety by treatment
bar_q26 <- custom_barplot(
  data = data, 
  x = q26, 
  group = group,
  subtitle = str_wrap(attr(data$q26, "label"), width = 85)
  )+
  theme(text = element_text(family = 'TeX Gyre Pagella'),
        legend.position = "right")

bar_q25
bar_q26

Briefly reviewing @fig-safety-bars, those in the treatment condition felt less concern over the potential for violence compared to those in the control. As for confidence in voter safety at election sites, it appears that more respondents in the treatment condition expressed that they were "Very confident" compared to those in the control, by approximately $8$ percentage points.

It would appear that the treatment vignette increased the likelihood that a respondent would select a higher response option compared to those in the control group. In order to assess whether that was the case, I conducted an ordered logistic regression on the relevant items using the experimental condition as a dichotomous predictor.

Those in the treatment condition had lower probability to express that they were 'Very' or 'Somewhat' concerned for violence in Maricopa County, AZ Elections, yet higher probability of expressing 'Not too' or 'Not at all' concerned. As for confidence in voter safety, moving from treatment to control, the difference in predicted probability of a respondent selecting "Very Confident" was $0.08$. @tbl-preds-diff1 and @tbl-preds-diff2 show the significant differences in predicted probabilities at 95$\%$ confidence^[The same models were run for items pertaining to safety concerns in one's local area but results of the treatment were not significantly different from the control. See Appendix D for full results table of ordinal logistic regression].

```{r}
#| label: ordinal-logit
#| include: false

# needed to duplicate `group` variable with different column name due to
# conflict with marginaleffects R package
data <- data |>
  mutate(condition = group)

# running ordinal logisitc regression models (proportional odds logistic regression)
m.q25 <- MASS::polr(q25 ~ condition, data = data, method = "logistic", Hess = T)
m.q26 <- MASS::polr(q26 ~ condition, data = data, method = "logistic", Hess = T)
m.q36 <- MASS::polr(q36 ~ condition, data = data, method = "logistic", Hess = T)
m.q37 <- MASS::polr(q37 ~ condition, data = data, method = "logistic", Hess = T)

# "Outcome predicted by a fitted model on a specified scale for a given
# combination of values of the predictor variables, such as their observed
# values, their means, or factor levels (a.k.a. "reference grid")".

# predictions(): unit-level (conditional) estimates.

# predicted probability estimates by treatment or control
q25_preds <- marginaleffects::predictions(m.q25, by = "condition", type = "probs")
q26_preds <- marginaleffects::predictions(m.q26, by = "condition", type = "probs")
q36_preds <- marginaleffects::predictions(m.q36, by = "condition", type = "probs")
q37_preds <- marginaleffects::predictions(m.q37, by = "condition", type = "probs")

# this provides the difference in predicted probabilities between treatment and
# control
pdiff_q25 <- marginaleffects::predictions(
  m.q25,
  by = "condition",
  type = "probs",
  hypothesis = ~ reference | group
)

pdiff_q26 <- marginaleffects::predictions(
  m.q26,
  by = "condition",
  type = "probs",
  hypothesis = ~ reference | group
)

```{r}
#| label: tbl-preds-diff1
#| tbl-colwidths: 80
#| tbl-cap: "Difference in Predicted Probabilities in Concern for violence, Threats of Violence, or Intimidation between Treatment and Control"

tab_q25_pdiff <- pdiff_q25 |>
  dplyr::as_tibble() |>
  dplyr::rename(
    Violence = group,
    z = statistic,
    'Pr(>|z|)' = p.value) |>
  select(Violence,
         hypothesis,
         estimate,
         std.error,
         z,
         'Pr(>|z|)',
         conf.low,
         conf.high) |>
  insight::format_table(stars = T, digits = 3)

tinytable::tt(
  tab_q25_pdiff, 
  notes = "Results derived from proportional odds logistic regression model. Estimates are differences in predicted probabilities transformed from log odds ratios, and represent difference in estimated prediction of choosing a particular response or higher") |> 
  tinytable::style_tt(
    fontsize = 0.80 # fontsize of text in table in em units (0.75*default)
    ) |>  
  tinytable::style_tt(i = "notes", fontsize = 0.75)

```{r}
#| label: tbl-preds-diff2
#| tbl-colwidths: 80
#| tbl-cap: "Difference in Predicted Probabilities in Confidence that voters will be safe to vote in-person between Treatment and Control"

tab_q26_pdiff <- pdiff_q26 |>
  dplyr::as_tibble() |>
  dplyr::rename('Voter Safety' = group,
                z = statistic,
                'Pr(>|z|)' = p.value) |>
  select('Voter Safety',
         hypothesis,
         estimate,
         std.error,
         z,
         'Pr(>|z|)',
         conf.low,
         conf.high) |>
  insight::format_table(stars = T, digits = 4)

tinytable::tt(
  tab_q26_pdiff, 
  notes = "Results derived from proportional odds logistic regression model. Estimates are differences in predicted probabilities transformed from log odds ratios, and represent difference in estimated prediction of choosing a particular response or higher") |>  
  tinytable::style_tt(
    fontsize = 0.80 # fontsize of text in table in em units (0.75*default)
    ) |>  
  tinytable::style_tt(i = "notes", fontsize = 0.75) # style font of table notes

@@

Publicized announcements that election officials are actively recruiting former military service members to work as election staff and volunteers is associated with greater confidence in elections compared to election worker recruitment announcements that do not explicitly mention targeted efforts to recruit former military service members. In addition, such public announcements eased concerns about violence and increased confidence that voters would be safe to vote in person at election sites. Of course, the higher levels of confidence in elections and voter safety observed among the treatment group were limited to considerations about elections held in Maricopa County, AZ---presumably beyond one's local area. This is to be expected bearing in mind that confidence in elections for one's local area is usually higher than for those beyond. Thus, there's already more insecurity concerning elections elsewhere for whatever reason.

One obvious contribution of these results shows that the perceived trustworthiness of elections administration is amendable to generalized impressions about the kind and character of the people who work and volunteer as election staff. Whatever election officials do to boost public confidence may be overshadowed by impressions about who is staffing election sites.

It should be noted that these results are limited to one's expectations in anticipation of election night. Whether the increase in confidence is sustained post-election night cannot be inferred from these results. Moreover, It is yet unclear whether actually encountering and interacting with veterans working as staff or volunteers at election sites does much to improve voter confidence after election night compared to anyone else. It may be the case that assumptions that veterans are filling spots as election workers and volunteers simply primes positive expectations as one heads into an election site, potentially improving voters perceptions of their own experience on the way out. Further research would be needed to test such an idea.

In general, public confidence appears strongest when the elections under consideration regard those within one's local community. Yet, such strong confidence is not generalized outward to the institution of elections administration as a whole. Insecurity in the integrity of the electoral process and its administration may be naturally higher for elections that occur further from home. Whether that is due to some local favorability bias or some outward negativity bias is hard to say. And seeing as how insecurity presents vulnerability, it is perhaps far easier to inspire a general sense of distrust in the process and its administration in places more abstract.

In the anticipatory period prior to national elections, publicized efforts to recruit veterans to work as election staff and volunteers may be a small, but positive, step towards reducing insecurity in elections that occur elsewhere. Especially among those who maintained a steadfast degree of distrust in elections. Despite a lack of evidence that systematic electoral fraud had occurred in the 2020 election, a substantial proportion of the population were unwilling to affirm the legitimacy of the results. However, despite such legitimacy beliefs, most were not consistent as they expressed varying levels of trust in different aspects of elections administration. These results say more about the influence that military veterans have among those who are most insecure than it does about the public more generally.

{{< pagebreak >}}

@@

{{< pagebreak >}}

```{r}
#| label: tbl-4
#| tbl-cap: "Survey Items and Response Options"
#| tbl-colwidths: [70, 30]

var_label_tab <- function(x){
    purrr::map(x, ~attr(., "label")) |>
    purrr::map(~ifelse(purrr::is_null(.), "No label", .)) |>
    tibble::enframe(name = "var", value = "var_label") |>
    tidyr::unnest(cols = c(var_label))
}

Questions <- data |> 
  dplyr::select(q7, q25, q26, q36, q37) |>
  dplyr::rename_with(.cols = everything(), .fn = ~snakecase::to_any_case(.,case = "upper_camel")) |> 
  var_label_tab() |>
  dplyr::select(Question = var_label) |> 
  dplyr::pull(Question)

items_responses <- data.frame(
  Questions = Questions,
  Response =c(
    "- Legitimate\n- Not legitimate",
    "- Very concerned\n- Somewhat concerned\n- Not too concerned\n- Not at all concerned",
    "- Not at all confident\n- Not too confident\n- Somewhat confident\n- Very confident",
    "- Very concerned\n- Somewhat concerned\n- Somewhat unconcerned\n- Not at all concerned",
    "- Not at all confident\n- Not too confident\n- Somewhat confident\n- Very confident")
)

# wrap in a minipage environment
minipagify <- function(x) {
  sprintf(
    "\\minipage{\\textwidth}%s\\endminipage",
    sapply(x, mark, "latex")
  )
}

# only in LaTeX
is_latex <- identical(knitr::pandoc_to(), "latex")

tab <- tinytable::tt(items_responses, width = c(0.7, 0.3)) |>
  tinytable::style_tt(j = c(1,2), fontsize = 0.8, align = "l") |>
  tinytable::format_tt(markdown = TRUE) |>
  tinytable::format_tt(j = 2, fn = if (is_latex) minipagify else NULL)

tab

@@

<!-- {{< pagebreak >}} -->

@@

```{r}
#| label: tbl-demog
#| tbl-cap: "Sample Demographics"
#| tbl-pos: 'H'

@@

@tbl-logit shows results of a logistic regression test of random assignment to the treatment group. Demographics such as age, gender, race, educational attainment, and party ID are included as predictor variables. Note that 19 missing observations were deleted. A $\chi^2 = 11.010$, with 15 degrees of freedom and associated p-value > 0.05 ($p = 0.75$) confirms that none of the demographic predictor variables significantly increased the log-odds---in turn, the probability---of being assigned to the treatment group.

@@

# simple logit of test of random assignment to treatment
# treatment is DV (treatment = 1, control = 0)
# demographic variables are IVs
m1 <- glm(group ~ age_4cat + gender_3cat + race2 + educ_4cat + partyid_3cat, 
          family = binomial(link = "logit"), data = data)

# use `gtsummary` package to build regression table, then output as tibble
logit_tab <- gtsummary::tbl_regression(
    m1,
    intercept = F,
    label = list(
      age_4cat ~"Age",
      gender_3cat ~ "Gender",
      race2 ~ "Race/Ethnicity",
      educ_4cat ~ "Education",
      partyid_3cat ~ "Party ID")) |>
  gtsummary::bold_labels() |>
  # gtsummary::show_header_names()
  gtsummary::modify_header(
    label = " ",
    stat_nevent = "**Treatment N**") |>
  # output as_tibble
  gtsummary::as_tibble() |>
  # omit p.values
  dplyr::select(-4)

# transform tibble dataframe into `tinytable` and format as markdown
# replace NA with blank " " (or "-" if desired)
logit_tab <- tinytable::tt(logit_tab, width = 0.8) |> 
  tinytable::format_tt(markdown = T) |> 
  tinytable::format_tt(replace = " ")

# create 2nd table containing only logit model fit statistics
# using custom function, get logit regression model fit stats... 
gof_stats <- lrm.gof.stats(m1) |> 
  # omit certain stats (saves space)
  dplyr::select(
    nobs, logLik, df, chisq, P..chi., deviance, null.deviance,
    -AIC, -BIC, -df.null, -df.residual) |> 
  # rename column variables
  dplyr::rename(
    "Num.Obs"       = "nobs",
    "Null Deviance" = "null.deviance",
    "Deviance"      = "deviance",
    "DF"            = "df",
    "Log.Lik"       = "logLik",
    "\\chi^2"       = "chisq",
    "P(>\\chi^2)"   = "P..chi."
  ) |> 
  # pivot into two column long-form dataframe
  tidyr::pivot_longer(cols = everything(),names_to = " ", values_to = '**log(OR)**') |>
  # pipe into tinytable
  tinytable::tt(width = 0.8) |> 
  # ensure rows with math text are identified and formatted as math (i.e., $..$)
  tinytable::format_tt(i = c(4,5), j = 1, math = T)

# using `rbind2` function, bind 2nd table at bottom of first table 
tinytable::rbind2(logit_tab, gof_stats, use_names = F, headers = F) |> 
  # add horizontal line just above fit stats
  tinytable::style_tt(i = 26, line = "t", line_color = "black", line_width = 0.1) |>
  # adjust fontsize of text within table
  tinytable::style_tt(fontsize = 0.8) |>
  # finally, format markdown and blank out any NA fillers
  tinytable::format_tt(markdown = T, replace = " ")

@@

{{< pagebreak >}}

@@

{{< pagebreak >}}

# Appendix D: Predicted Probabilities of Safety Concerns by Treatment Condition {.appendix}
